---
title: "EDA"
output: html_document
---

Here, we will explore the data collected during the Spring 2019 flowering field season at the Granite Mountains desert research center. 

Some key datasets:
+ Bird visitation at focal cactus: the number of bird visitations to a focal cactus individual, with manipulated showiness of flowers (both real and fake) on a cactus at zero, fifteen, and thirty flowers per cactus. 
+ Cactus allocation: a measurement of reproductive and non reproductive structures of 100 cactus indivduals. This dataset will be paired with data on fruit production of the same individuals.
+ Line transects: site-wide measurements of bird diversity, mesohabitats, and behaviors
+ Cactus architecture temperature: records of micro-habitat temperature related to cacti by measuring the top of the cactus's canopy, within the cactus's canopy, and the cactus's understory.
+ Interspecific allocation of cacti: haphazardly collected data on the reproductive output as well as size of three cactus species (*Opuntia basilaris*, *Cylindropuntia echinocarpa*, and *Cylindropuntia acanthocarpa*)

```{r libraries, include=FALSE}
library(ggplot2)
library(dplyr)
library(kableExtra)
```

```{r datasets, include=FALSE}
#bring in datasets of interest
site <- read.csv("~/Masters/Flower_CactusBirdInteractionR/data/line_transects/line_transects.csv")
head(site)

```

```{r Site Wide}
#histogram of bird behaviors
ggplot(site, aes(x=behavior)) + geom_bar() + theme(axis.text.x = element_text(angle = 60, hjust = 1))
#lump into smaller groups (e.g. movement, passive, vocalization, mutualistic interaction)

#histogram of bird mesohabitats
ggplot(site, aes(x=mesohabitat)) + geom_bar() + theme(axis.text.x = element_text(angle = 60, hjust = 1))

#histogram of bird species
ggplot(site, aes(x=taxa)) + geom_bar() + theme(axis.text.x = element_text(angle = 60, hjust = 1))
#lump into broader taxa and functional groups

#histogram of bird distance from transect
ggplot(site, aes(x=distance)) + geom_bar() 


```

